---
title: "R/Python Package Download Analysis"
subtitle: "Posit Demo Dataset Summary"
author: "Generated for Demonstration Purposes"
date: today
format: 
  html:
    toc: true
    code-fold: true
    theme: cosmo
execute:
  warning: false
  message: false
---

## Overview

This report analyzes synthetic R and Python package download data, demonstrating typical data analysis workflows and highlighting data quality issues that require attention.

**Note: This dataset is completely synthetic and generated for demonstration purposes.**

## Data Loading and Initial Exploration

```{r setup}
library(tidyverse)
library(plotly)
library(DT)
library(lubridate)

# Load the data
downloads <- read_csv("data.csv")

# Display basic information
cat("Dataset dimensions:", nrow(downloads), "rows ×", ncol(downloads), "columns\n")
```

### Dataset Structure

```{r structure}
# Display structure and summary
glimpse(downloads)
```

```{r summary-table}
# Create a nice summary table
downloads %>%
  summary() %>%
  knitr::kable(caption = "Dataset Summary Statistics")
```

## Data Quality Issues 🚨

Let's identify the obvious data quality problems in this dataset:

### Negative Download Counts

```{r negative-downloads}
negative_downloads <- downloads %>%
  filter(daily_downloads < 0)

cat("Found", nrow(negative_downloads), "records with negative download counts:\n")

negative_downloads %>%
  arrange(daily_downloads) %>%
  head(10) %>%
  datatable(caption = "Records with Negative Downloads") %>%
  formatStyle("daily_downloads", 
              backgroundColor = styleInterval(0, c("lightcoral", "white")))
```

### Extreme Outliers

```{r extreme-outliers}
# Identify extremely high download counts (likely errors)
outliers <- downloads %>%
  filter(daily_downloads > 1000000)  # More than 1 million downloads per day

cat("Found", nrow(outliers), "records with suspiciously high download counts:\n")

outliers %>%
  arrange(desc(daily_downloads)) %>%
  datatable(caption = "Records with Extreme Download Counts") %>%
  formatStyle("daily_downloads",
              backgroundColor = styleInterval(1000000, c("white", "orange")))
```

### Zero Downloads for Popular Packages

```{r zero-downloads}
popular_packages <- c("ggplot2", "pandas", "numpy", "tidyverse", "matplotlib")

zero_popular <- downloads %>%
  filter(package_name %in% popular_packages & daily_downloads == 0)

cat("Found", nrow(zero_popular), "records where popular packages have zero downloads:\n")

zero_popular %>%
  datatable(caption = "Popular Packages with Zero Downloads") %>%
  formatStyle("daily_downloads",
              backgroundColor = styleInterval(0.5, c("lightyellow", "white")))
```

## Visualization Analysis

### Download Trends Over Time

```{r downloads-over-time}
# Clean data for visualization (remove obvious errors)
clean_downloads <- downloads %>%
  filter(daily_downloads >= 0 & daily_downloads <= 100000)

monthly_trends <- clean_downloads %>%
  mutate(month = floor_date(download_date, "month")) %>%
  group_by(month, package_type) %>%
  summarise(avg_downloads = mean(daily_downloads, na.rm = TRUE), .groups = "drop")

p1 <- ggplot(monthly_trends, aes(x = month, y = avg_downloads, color = package_type)) +
  geom_line(size = 1.2) +
  geom_point(size = 2) +
  labs(
    title = "Average Monthly Downloads by Package Type",
    subtitle = "Data cleaned to remove obvious errors",
    x = "Month",
    y = "Average Daily Downloads",
    color = "Package Type"
  ) +
  theme_minimal() +
  scale_y_continuous(labels = scales::comma_format())

ggplotly(p1)
```

### Package Type Distribution

```{r package-distribution}
package_summary <- clean_downloads %>%
  group_by(package_type) %>%
  summarise(
    count = n(),
    avg_downloads = round(mean(daily_downloads)),
    median_downloads = median(daily_downloads),
    total_downloads = sum(daily_downloads),
    .groups = "drop"
  )

package_summary %>%
  datatable(caption = "Package Type Summary (Cleaned Data)") %>%
  formatCurrency(c("avg_downloads", "median_downloads", "total_downloads"), 
                  currency = "", digits = 0)
```

### Top Packages by Downloads

```{r top-packages}
top_packages <- clean_downloads %>%
  group_by(package_name, package_type) %>%
  summarise(
    total_downloads = sum(daily_downloads),
    avg_rating = round(mean(user_rating), 1),
    .groups = "drop"
  ) %>%
  arrange(desc(total_downloads)) %>%
  head(15)

p2 <- ggplot(top_packages, aes(x = reorder(package_name, total_downloads), 
                               y = total_downloads, 
                               fill = package_type)) +
  geom_col() +
  coord_flip() +
  labs(
    title = "Top 15 Packages by Total Downloads",
    subtitle = "Based on cleaned data",
    x = "Package Name",
    y = "Total Downloads",
    fill = "Package Type"
  ) +
  theme_minimal() +
  scale_y_continuous(labels = scales::comma_format())

ggplotly(p2)
```

## Data Cleaning Recommendations

Based on this analysis, the following data cleaning steps are recommended:

1. **Remove or investigate negative download counts** - These are impossible and likely data entry errors
2. **Cap or investigate extreme outliers** - Downloads > 1M per day for individual packages seem unrealistic
3. **Investigate zero downloads for popular packages** - May indicate missing data or system issues
4. **Validate data collection processes** - To prevent similar issues in future datasets

## Interactive Data Explorer

```{r interactive-table}
downloads %>%
  datatable(
    caption = "Complete Dataset (Including Errors for Demo)",
    filter = "top",
    options = list(pageLength = 25, scrollX = TRUE)
  ) %>%
  formatStyle(
    "daily_downloads",
    backgroundColor = styleInterval(c(0, 100000), 
                                   c("lightcoral", "white", "orange"))
  )
```

---

*This report demonstrates typical data quality issues and analysis workflows using synthetic data generated for Posit demonstrations.*